A Constraint Solving Approach to Parikh Images of Regular Languages

### 2023-07-11 example.tex
more refs

In this section we will introduce an intuition for how a string constraint
problem in an automata-based solver like \OstrichPlus{} is translated to a
Parikh automata intersection problem and solved using \Calculus{}.

We use PCRE regular expression notation here and throughout the paper, writing
them $\Regex{\texttt{like this}}$. This means that $\RegexOr{}$ is alternation,
$\mathtt{*}$ the Kleene star, and $\mathtt{.}$, matches any single character.
For the length of a string $s$, we write $\Length{s}$. All number variables in
this example are in $\Naturals$ with $0$.

\begin{example}\label{ex:string-constraints} Consider the following set of
    string constraints:
\begin{constraints}
    \item\label{const:s1-in-c-dd} $s_1 \in \Language(\Regex{c\RegexOr{}dd})$.
    \item\label{const:s2-in-b} $s_2 \in \Language(B)$, the language accepted by
    automaton~$B$ of \cref{fig:aut_b}.
    \item\label{const:s1-substring} $s_1 = \Substring(s_2, i, n)$, that is $s_1$ is an
    $n$-length substring of $s_2$ starting at offset $i$.
    \item\label{const:more-inside-than-before} $n > i$, that is what comes
    before $s_1$ in $s_2$ is shorter than $s_1$.
    \item\label{const:something-before-and-after} $i > 0 \land \Length{s_2} -i -n > 0$, that
    is there is at least one character in $s_2$ after $s_1$.
\end{constraints}
\end{example}

To solve these constraints, we will first perform translation into Parikh
automata, then break down how the constraints will be handled by \Calculus{}.
This example will recur in a more formal fashion in \cref{sec:generalised},
which can be read in parallel if (even) more detail is desired.

>>>>

In this section, we will introduce an intuition for how a string constraint
problem in an automata-based solver like \OstrichPlus{} is translated to a
Parikh automata intersection problem, and solved using \Calculus{}.

We use PCRE regular expression notation here and throughout the paper, writing
them $\Regex{\texttt{like this}}$. This means that $\Regex{\RegexOr{}}$ is alternation,
$\Regex{\mathtt{*}}$ the Kleene star, and $\Regex{\mathtt{.}}$ matches any single character.
For the length of a string $s$, we write $\Length{s}$. All number variables in
this example are in $\Naturals$ with $0$.

\begin{example}\label{ex:string-constraints} Consider the following set of
    string constraints over string variables~$s_1, s_2$ and integer
    variables~$n, i$:
\begin{constraints}
    \item\label{const:s1-in-c-dd} $s_1 \in \Language(\Regex{c\RegexOr{}dd})$.
    \item\label{const:s2-in-b} $s_2 \in \Language(B)$, the language accepted by
    automaton~$B$ of \cref{fig:aut_b}.
    \item\label{const:s1-substring} $s_1 = \Substring(s_2, i, n)$, that is $s_1$ is an
    $n$-length substring of $s_2$ starting at offset $i$.
    \item\label{const:more-inside-than-before} $n > i$, that is what comes
    before $s_1$ in $s_2$ is shorter than $s_1$.
    \item\label{const:something-before-and-after} $i > 0 \land \Length{s_2} -i -n > 0$, that
    is there is at least one character in $s_2$ before and after $s_1$.
\end{constraints}
\end{example}

Although the constraints are simple, it should be noted that
decidability results for string constraints involving integers are
notoriously hard to obtain; most state-of-the-art string solvers, for
instance Z3~\cite{Z3}, different versions of
Z3-str~\cite{Z3-str,DBLP:conf/fm/MoraBKNG21}, or
cvc5~\cite{cvc5}
are not guaranteed to be complete or to terminate on constraints of this kind.

A decision procedure for constraints of this kind is defined in
\cite{ostrich-plus}.
%
To solve the constraints, we will first perform translation into Parikh
automata, then break down how the constraints will be handled by \Calculus{}.
This example will recur in a more formal fashion in \cref{sec:generalised},
which can be read in parallel if (even) more detail is desired.
